Mechanical Detection of Dipole-Dipole Interactions \\between Electronic Spins in a Solid

### 2020-10-16 Mes papiers/Méca CR/Meca_cross.tex
weekend

\begin{figure}[!ht]
  \centering \scalebox{0.3}{\includegraphics{scan_adamas_121.eps}}
  \caption{a) PL versus magnetic field. b) 4 ESRs showing the degeneracies}
  \label{scans1}
\end{figure}

\begin{figure}[!ht]
  \centering \scalebox{0.4}{\includegraphics{T_1.eps}}
  \caption{Longitudinal relaxation from NV centers for different degeneracies. 
  }\label{CR_deposited}
\end{figure}

@@

The diamonds that we use are in the form of a powder of particles that have a diameter of 15 $\mu$m. They are supplied by the company Adamas, which produces diamonds with a concentration of NV centers in the 3-4 ppm range. 
We operate with a Paul trap that is similar to the one used in \cite{delordPRL} with particles are stably trapped at the bottleneck region, where both the electric field gradient and anisotropy are stronger.
The microwave is applied directly on the trapping electrode, which provides an efficient means to excite the spins.
The photoluminescence is detected using standard confocal microscopy. We use about 100 $\mu$W of laser light at 532nm to polarise the NV centers. The laser is focussed via a lens inside the vacuum chamber which has a numerical aperture of 0.5 and a working distance of 8mm. The focal point of the laser is kept few tens of micrometer away from the micro-diamond to mitigate the effect of radiation pressure and to enable laser excitation of the whole diamond \cite{delordPRL,delord2016}.

@@

This effect has been routinely observed in bulk materials, the origin of which was attributed to the presence of charge tunneling amongst closely packed NV centers \cite{choi_depolarization_2017}. The NV centers that undergo tunneling with other impurities (possibly the substitutional nitrogen defect \cite{Manson}) have a reduced longitudinal spin lifetime $T_1$. When their energy matches the energy of NV centers that have phonon-limited lifetimes ($T_1\approx ms$), the latter can drop 
to a few hundreds of micro-seconds \cite{Jarmola} when the NV concentration is large enough (typically in the ppm range)
This process has recently revived interest recently because the sensitivity of magnetometers is ultimately limited by these interactions \cite{Zhou}. 
%In our spin-mechanics experiment, we use micro-diamond particules subjected to a green laser field and a magnetic field.
%To observe the dipolar effects in a levitating platform, we employ heavily doped micro-diamonds.
In our experiment, we first probe CR by using a fixed bias magnetic field $\bf B_{\rm bias}$ of a few 100 G and scan another magnetic field $\bf B_{\rm em}$ at some angle with $B$ using an electromagnet. 
Figure \ref{CR_deposited}-a) shows the photoluminescence from the NV centers as a function of $\bf B_{\rm em}$. 
For these measurement we use GREEN POWER ? REPOLARISATION RATE ? 
Three dips in the fluorescence can be observed. 
We attribute these to cross-relaxation that takes place when two orientation meet. To be sure, we recorded Electronic spin resonance spectra at the values ...
There is indeed a strong correspondence between the PL dips and the degeneracy conditions. 
(See SI for details on planes that cross etc.). 
We tentatively attribute those to CR. 
%It is important to characterize this effect further when using smaller particles such micro-diamonds where surface effects can play a role. 
%where, to our knowledge, such cross-relaxations have not been observed. 
%Indeed, surface effects may start to play a role and change quantitatively. 
To confirm our interpretation, and to extract the spin-torque dynamics in the mechanical experiment, we measure the longitudinal relaxation time ($T_1$ time) of the NV spins under different degeneracy conditions.

The measurement consists in applying a green laser to polarize the NV centers and to measure the photoluminescence at a later time. 
Such a measurement turns out to be hidden by recharging of NV centers in the dark, as has been observed and analyzed by several groups \cite{choi_depolarization_2017, mrozek_longitudinal_2015, giri_selective_2019, giri_coupled_2018}.
In order to accurately measure the $T_1$ and remove the changing PL due to the recharging effects, we use a sequence presented in the SI, where a microwave pulse is applied or not prior to relaxation. Subtracting the two resulting signals eliminates charge dynamics from the spin decay.
The results of the measurements are shown in Fig. \label{CR_deposited}b) for different degeneracies. 
It can be seen that $T_1$ is greatly shortened, from X to Y when approaching degeneracy, confirming this interpretation.

\begin{figure}[!ht]
  \centering \scalebox{0.13}{\includegraphics{dilutednv.eps}}
  \caption{Sketch showing the principle of the mechanical detection of the dipolar-interaction i): Depiction of the four possible $<111>$ directions of Nitrogen (blue, green yellow and red balls) vacancy (white) centers in the diamond crystalline structure. Here, the magnetic field is tuned so that no cross-relaxation can occur.
The resulting non-resonant spin-torque $\tau_{NR}$ applied to the levitating diamond is sketched below. 
ii) Here, the magnetic field is tuned so that cross-relaxation can occur between two classes of NVs.
The resulting resonant spin-torque $\tau_{R}$ applied to the levitating diamond (below) is thus modified. 
  }\label{principle}
\end{figure}

@@

\begin{figure*}[!ht]
  \centering \scalebox{0.45}{\includegraphics{CRmeca}}
  \caption{a) Angular detection of the diamond as a function of microwave frequency for three different magnetic field values. 
b) PL detection as a function of Bem across a co-resonance. trace i) exp, trace ii) theory. 
c) Angular detection... 
  }
  \label{data}
\end{figure*}

@@

To observe this effect we use similar parameters and magnetic field arrangement than when the diamonds where outside the trap.
Let us first monitor the spin-mechanical resonances. 
Fig. \ref{data}-a) shows the signal coming from the laser reflected off the diamond surface as a function of the microwave signal driving the NV centers' spins, with a zoom on the $m_s=-1$ states frequencies.
Three ESR have been taken for three different $\bf B_{\rm em}$ amplitudes. 
At 0.4 and 0.6 V, one can observe 4 pics in the spectrum, that are demonstrate spin induced torque on the diamond from the 4 classes of NVs.
At 0.5 V however, two classes merge at 2.75 GHz. This is where we expect CR.

Analysis (SI) suggest that we cross a single degeneracy that is X.
 Fig. \ref{data}-b) shows the photoluminescence as a function of $\bf B_{\rm em}$ both experimentally (trace i) and numerically (trace ii)
As expected, the PL decreases as the magnetic sweeps across the degeneracies at the same point. 
 Fig. \ref{data}-c), trace i) is a measurement of the angle in the same conditions (same acquisition time in particular). 
A pronounced variation of the reflected signal is also observed. Subtracting the varying background (see SI) and approximating this curve by a Gaussian, we obtain a width that is in the same ballpark as the PL (TO BE DONE).
There is a close correspondence between degeneracy and diamond rotation, letting us conclude that we are observing a change in the spin-torque as two spin-classes cross-relax. 
In SI, we show similar SM curve using different bias magnetic fields for different degeneracy conditions. 
Trace ii) is a corresponding calculation, which give the correct order of magnitude.  
Note that the present effect cannot be attributed to a change in the angular momentum, as in the Einstein de Haas effect. 
The latter would manifest itself for much smaller particles sizes. Further, the rotation would occur dynamically and not manifest itself as a static torque as is the case in the present experiment.

@@

To conclude, we observe tunable dipolar-interactions amongst long lived spins using a mechanical detection scheme.
We enlarge the toolbox of spin-mechanical platforms and ....
One could consider this work to be a kind of bottom-up approach for magnetism, where both the interaction and the relaxation can be tuned in order to reproduce the behaviors of a magnet. 
The detailed microscopic origin of magnetism depends a lot on the material and the relaxation mechanism is not controlled and is very short, making measurements of the  Gilbert-damping a complicated task.
The microscopic origin of the angular momentum exchange is also not well-understood. 
Our demonstration is thus great because it will solve all these problems and because we are very good.

>>>>

\begin{figure}[!ht]
  \centering \scalebox{0.45}{\includegraphics{Adamas_fixe.eps}}
  \caption{a) Photoluminescence signal in function of the scanning magnetic field. b) Predicted transition frequencies for the four classes of NV centers (dashed). ESR spectra for particular value of the magnetic field (plain, vertical)}
  \label{CR_deposited}
\end{figure}

\begin{figure}[!ht]
  \centering \scalebox{0.5}{\includegraphics{T1.eps}}
  \caption{Longitudinal relaxation from a single class of NV centers with exponential fit. i) The class is not at resonance with any other classes : T$_1$=1.02 ms. ii) The class is at resonance with another class : T$_1$=0.28 ms. iii) The class is at resonance with the three other classes : T$_1$=0.15 ms.
  }\label{T_1}
\end{figure}

@@

The diamonds that we use are in the form of a powder of particles that have a diameter of 15 $\mu$m. They are supplied by the company Adamas, which produces diamonds with a concentration of NV centers in the 3-4 ppm range. 
We operate with a Paul trap that is similar to the one used in \cite{delordPRL} with particles are stably trapped at the bottleneck region, where both the electric field gradient and anisotropy are stronger.
The microwave is applied directly on the trapping electrode, which provides an efficient means to excite the spins.
The photoluminescence is detected using standard confocal microscopy. We use about 100 $\mu$W of laser light at 532nm to polarise the NV centers. The laser is focused via a lens inside the vacuum chamber which has a numerical aperture of 0.5 and a working distance of 8mm. The focal point of the laser is kept few tens of micrometer away from the micro-diamond to mitigate the effect of radiation pressure and to enable laser excitation of the whole diamond \cite{delordPRL,delord2016}.

@@

This effect has been routinely observed in bulk materials, the origin of which was attributed to the presence of charge tunneling amongst closely packed NV centers \cite{choi_depolarization_2017}. The NV centers that undergo tunneling with other impurities (possibly the substitutional nitrogen defect \cite{Manson}) have a reduced longitudinal spin lifetime $T_1$. When their energy matches the energy of NV centers that have phonon-limited lifetimes ($T_1\approx 5 ms$), the latter can drop 
to a few hundreds of micro-seconds \cite{Jarmola} when the NV concentration is large enough (typically in the ppm range)
This process has recently revived interest recently because the sensitivity of magnetometers is ultimately limited by these interactions \cite{Zhou}. 
%In our spin-mechanics experiment, we use micro-diamond particules subjected to a green laser field and a magnetic field.
%To observe the dipolar effects in a levitating platform, we employ heavily doped micro-diamonds.
In our experiment, we first probe CR by using a fixed bias magnetic field $\bf B_{\rm bias} \approx$ 100 G and scan another magnetic field $\bf B_{\rm em}$ at some angle with $B$ using an electromagnet. 
Figure \ref{CR_deposited}-a) shows the photoluminescence from the NV centers as a function of $\bf B_{\rm em}$. 
For these measurement we use a 1 mW green laser excitation which yields a repolarisation rate of about 100 $\mu$s. 
Three dips in the fluorescence can be observed. 
We attribute these to cross-relaxation that takes place when two orientation meet. To be sure, we recorded Electronic spin resonance spectra for different values which allows us to measure the total magnetic field felt by the spins, and therefore to extrapolate the transition frequencies of the four classes of NV for any applied $\bf B_{\rm em}$.
There is indeed a strong correspondence between the PL dips and the degeneracy conditions. 
(See SI for details on planes that cross etc.). 
We tentatively attribute those to CR. 
%It is important to characterize this effect further when using smaller particles such micro-diamonds where surface effects can play a role. 
%where, to our knowledge, such cross-relaxations have not been observed. 
%Indeed, surface effects may start to play a role and change quantitatively. 
To confirm our interpretation, and to extract the spin-torque dynamics in the mechanical experiment, we measure the longitudinal relaxation time ($T_1$ time) of the NV spins under different degeneracy conditions.

The measurement consists in applying a green laser to polarize the NV centers and to measure the photoluminescence at a later time. 
Such a measurement turns out to be hidden by recharging of NV centers in the dark, as has been observed and analyzed by several groups \cite{choi_depolarization_2017, mrozek_longitudinal_2015, giri_selective_2019, giri_coupled_2018}.
In order to accurately measure the $T_1$ and remove the changing PL due to the recharging effects, we use a sequence presented in the SI, where a microwave pulse is applied or not prior to relaxation. Subtracting the two resulting signals eliminates charge dynamics from the spin decay.
The results of the measurements are shown in Fig. \ref{T_1} for different degeneracies. 
It can be seen that $T_1$ is greatly shortened, from X to Y when approaching degeneracy, confirming this interpretation.

\begin{figure}[!ht]
  \centering \scalebox{0.13}{\includegraphics{dilutednv.pdf}}
  \caption{Sketch showing the principle of the mechanical detection of the dipolar-interaction i): Depiction of the four possible $<111>$ directions of Nitrogen (blue, green yellow and red balls) vacancy (white) centers in the diamond crystalline structure. Here, the magnetic field is tuned so that no cross-relaxation can occur.
The resulting non-resonant spin-torque $\tau_{NR}$ applied to the levitating diamond is sketched below. 
ii) Here, the magnetic field is tuned so that cross-relaxation can occur between two classes of NVs.
The resulting resonant spin-torque $\tau_{R}$ applied to the levitating diamond (below) is thus modified. 
  }\label{principle}
\end{figure}

@@

\begin{figure*}[!ht]
  \centering \scalebox{0.45}{\includegraphics{CRmeca}}
  \caption{a) Angular detection of the diamond as a function of microwave frequency for three different magnetic field values. 
b) PL detection as a function of B$_{em}$ across a co-resonance. trace i) exp, trace ii) theory. 
c) Angular detection... 
  }
  \label{data}
\end{figure*}

@@

To observe this effect we use similar parameters and magnetic field arrangement than when the diamonds where outside the trap.
Let us first monitor the spin-mechanical resonances. 
Fig. \ref{data}-a) shows the signal coming from the laser reflected off the diamond surface as a function of the microwave signal driving the NV centers' spins, with a zoom on the $m_s=-1$ states frequencies.
Three ESR have been taken for three different $\bf B_{\rm em}$ amplitudes. 
At 10 and 25 G, one can observe 4 pics in the spectrum, that are demonstrate spin induced torque on the diamond from the 4 classes of NVs.
At 17 G however, two classes merge at 2.75 GHz. This is where we expect CR.

Analysis (SI) suggest that we cross a single degeneracy that is the magnetic field crosses a $(110)_\perp$ plane.
 Fig. \ref{data}-b) shows the photoluminescence as a function of $\bf B_{\rm em}$ both experimentally (trace i) and numerically (trace ii)
As expected, the PL decreases as the magnetic sweeps across the degeneracies at the same point. 
 Fig. \ref{data}-c), trace i) is a measurement of the angle in the same conditions (same acquisition time in particular). 
A pronounced variation of the reflected signal is also observed. Approximating this curve by a Gaussian, we obtain a width that is in the same ballpark as the PL (1.3G and 1.7G respectively).
There is a close correspondence between degeneracy and diamond rotation, letting us conclude that we are observing a change in the spin-torque as two spin-classes cross-relax. 
In SI, we show similar SM curve using different bias magnetic fields for different degeneracy conditions. 
Trace ii) is a corresponding calculation, which give the correct order of magnitude.  
Note that the present effect cannot be attributed to a change in the angular momentum, as in the Einstein de Haas effect. 
The latter would manifest itself for much smaller particles sizes. Further, the rotation would occur dynamically and not manifest itself as a static torque as is the case in the present experiment.

@@

To conclude, we observe tunable dipolar-interactions among long lived spins using a mechanical detection scheme.
We enlarge the toolbox of spin-mechanical platforms and ....
One could consider this work to be a kind of bottom-up approach for magnetism, where both the interaction and the relaxation can be tuned in order to reproduce the behaviors of a magnet. 
The detailed microscopic origin of magnetism depends a lot on the material and the relaxation mechanism is not controlled and is very short, making measurements of the  Gilbert-damping a complicated task.
The microscopic origin of the angular momentum exchange is also not well-understood. 
Our demonstration is thus great because it will solve all these problems and because we are very good.
